Learning Actionable Analytics (with applications for reducing defects and reducing runtimes)

### 2015-05-03 how.tex
Updates from ShareLaTeX

@@

Turning now to the {\em test} data, this is based to the {\em predictor}
to generate a baseline {\em before} performance score.
Note that if our {\em predictors} fail to perform adequately on the test data,
then we cannot trust them to comment on our plans. Accordingly,
if that performance is unsatisfactory, we abort.
Else, we (1)~apply the {\em planner} to alter the {\em test} data;
then (2)~apply the {\em predictor} to the altered data {\em test'};
then (3)~return the performance delta $\Delta$.

The clear advantage of this approach is that it can be readily and automatically
applied to domains with historical data. That said, the drawback with this approach
is that our assessment of the plans is only as good as the predictor. To 
mitigate for this {\em problem of poor planners}, it is wise to demand high standards form the {\em
satisfactory} threshold used in this \fig{frame}.

@@

This study explores data sets from two types of domains:
\bi
\item
In one type, we 
build plans to change static code attributes in order to reduce defects in
Java classes.  Here,,
performance can be measured in terms of recall (percent of faulty classes in
the test data detected
by the {\em predictor}). Using engineering judgement, we define satisfactory recall as anything less than 70\%. 
\item 
In the other type, we   build
plans to change configurations settings within a Makefile such that the resulting compiled
code will run faster. Here, performance can be measured in terms of difference
between the predicted runtime $p$ of test case items and their actual runtimes $a$
using  $s= 1- \frac{abs(a - p)}{a}$.
Using engineering judgement, we define satisfactory $s$ as anything less than 70\%. 
\ei
This paper explores 12 static code data sets
Ant, Camel, Ivy, Jedit, JBeans, Log4, Lucene, Poi, Synapse, Velocity, Xalan, Xerces\footnote{Available from the object-oriented defects section of the PROMISE respository openscience.us/repo/defect/ck.} and five configuration data sets Berkeley DB, Apache, SQLite, LLVM,
  x264\footnote{Available from the performance prediction section of PROMISE
  http://openscience.us/repo/performance-predict.}. For each of these data sets,
  we generated 25 {\em before} scores from
  random split train:test splits in the ratio 4:1 train.
 The median {\em before} recalls from  the static code data sets was:
%issue24
\[\mathrm{recall}=\{98,88,86,84,82,81,79,78,78,\overbrace{55,23,20}^{\mathrm{JBeans,Poi,Xalan}}\}\]
The median {\em before} $e$  scores from  the configuration data was
%issue 24
\[s=\{98,97,95,88,92\}\]
Note that we need to abort planning in only $3/17$ of our data sets.

This paper also explores five configuration data sets Berkeley DB, Apache, SQLite, LLVM,
  x264. The median {\em before} scores from these data sets (generated in 
-- which means we cannot generate plans for $3/12$ of our data sets. However, as shown
below, we were able to generate useful plans for the other $9/12$ of the data.

In the other type,

then studied to generate a prediction
on the test data. If the performance on that test data is not
satisfactory, thempredictions:
a {\em before} prediction (on the raw test data); and  an {\em after}
prediction (that is  applying the policies learned from the training data).

The improvement delta $\Delta$ is then computed by dividing {\em after} and {\em before}

This paper presents a novel policy generator called ``HOW''.
HOW is a near-linear time  instance-based
local search algorithm that looks for improvement trends in the data,
then adjusting examples to nudge them along that trend.
The experiments of this paper apply HOW to learning policies for 
software defect reduction and for compiling software that runs faster.
At least for the data explored here, HOW's reductions in defects and runtimes
are very large-- up to 50\% fewer defects and up to 50\% faster runtimes.

The rest of this paper is structured as follows.

>>>>

\subsection{Proposed Framework}

@@

Turning now to the {\em test} data, this is based to the {\em predictor}
to generate a baseline {\em before} performance score.
Note that if our {\em predictors} fail to perform adequately on the test data,
then we cannot trust them to comment on our plans. Accordingly,
if that performance is unsatisfactory, we abort.
Else, we (1)~apply the {\em planner} to alter the {\em test} data;
then (2)~apply the {\em predictor} to the altered data {\em test'};
then (3)~return the performance delta.

\subsection{When Not to Plan: Problem of Poor Planner}

The clear advantage of this approach is that it can be readily and automatically
applied to domains with historical data. That said, the drawback with this approach
is that our assessment of the plans is only as good as the predictor. To 
mitigate for this {\em problem of poor planners}, it is wise to demand high standards form the {\em
satisfactory} threshold used in this \fig{work}.

@@

How prevalent is the problem of poor planners? While
this is hard to answer in the general case, we can say that for the 17 data
sets explored in this study, only 3 suffer from this problem.
This study's    data   from two  types of domains:
\bi
\item
In one type, we 
build plans to change static code attributes in order to reduce defects in
Java classes.  Here,,
performance can be measured in terms of recall (percent of faulty classes in
the test data detected
by the {\em predictor}). Using engineering judgement, we define satisfactory recall as anything over   70\%. 
\item 
In the other type, we   build
plans to change configurations settings within a Makefile such that the resulting compiled
code will run faster. Here, performance can be measured in terms of difference
between the predicted runtime $p$ of test case items and their actual runtimes $a$
using  $s= 1- \frac{abs(a - p)}{a}$.
Using engineering judgement, we define satisfactory $s$ as anything over   70\%. 
\ei
This paper explores 12 static code data sets
Ant, Camel, Ivy, Jedit, JBeans, Log4, Lucene, Poi, Synapse, Velocity, Xalan, Xerces\footnote{Available from the object-oriented defects section of the PROMISE respository openscience.us/repo/defect/ck.} and five configuration data sets Berkeley DB, Apache, SQLite, LLVM,
  x264\footnote{Available from the performance prediction section of PROMISE
  http://openscience.us/repo/performance-predict.}. For each of these data sets,
  we generated 25 {\em before} scores from
  random split train:test splits in the ratio 4:1 train.
 The median {\em before} scores from  the static code data sets were:
%issue24
\[\mathrm{recall}=\{98,88,86,84,82,81,79,78,78,\overbrace{55,23,20}^{\mathrm{JBeans,Poi,Xalan}}\}\]
The median {\em before}   scores from  the configuration data were:
%issue 24
\[s=\{98,97,95,88,92\}\]
Note that we need to abort planning in only $3/17$ of our data sets.

\subsection{Testing for Grainer Causality}
With one small variation,  the above framework can be used to check if the learned
plans can be said to {\em cause} a change in the domain being explored.

In the literature, there are two main definitions of causality: Hill's H-causality
and Grainer's G-casualty. 
Of the two, Hill's definition is the most strict. An effect
is H-causal if it satisfies nine criteria including:
\be
\item
Plausibility;
\item
Strength of the effect; 
\item
Specificity: locality of the effect
to specialized groups;
\item
Support by experimental evidence;
\item
Coherence between cause-effects noted
in the field and in laboratory samples;
\item
Temporally: the effect has to follow the cause;
\item
Analogy: similar causes have similar effects;
\item
Biological gradient: more exposure to the cause leads to more of an
effect)
\item
Consistency: the same cause-effect pairings have been seen by different persons
in different places.
\ee
It is hard to demonstrate H-causality in software. For example, Hill's criteria emphasises repeatability and laboratory studies-- both of which are
hard to perform in software engineering (e.g. see the continued debate about
the value of using undergraduate students to surrogates for industrial
practitioners in software engineering research~\cite{Carver2003}). Also, Hill's requirement
for consistency is easier to achieve in studies of small closed systems (e.g. bacteria in a petri dish)
than in large socio-technical systems (e.g. open source projects). As Lofti Zadeh
commented: ``as complexity rises, precise statements lose meaning and meaningful statements lose precision''~\cite{zadeh.

Granger causality is a statistical concept of causality that is based on prediction. According to Granger causality, if a signal X1 "Granger-causes" (or "G-causes") a signal X2, then past values of X1 should contain information that helps predict X2 above and beyond the information contained in past values of X2 alone. Its mathematical formulation is based on linear regression modeling of stochastic processes (Granger 1969).

Hill1965

\cite{Seth2007,granger80}
